---
title: "Functional characteristics of vascular-inhabiting bacterial communities."
author: "Liyu Zhang"
date: "2021/12/31"
output: html_document
---

## a.Heatmap 
a, Heatmap of metabolic and ecological functions of bacteria based on FAPROTAX. Data are based on OTUs occurrence (number of OTUs capable of each function) in soil and plant compartments (log(x+1) transformed).  
```{r heatmap, warning=FALSE}
fap = read.table("faprotax.txt", header=T, row.names=1, sep="\t")
library(reshape2)
fap$fap=rownames(fap)
fap_melt <- melt(fap,id.vars = c("fap"),variable.name = "SampleID",value.name = "RA")
group = read.delim("metadata.txt", header=T, sep="\t",stringsAsFactors = FALSE,check.names = FALSE)
fap_merge <-merge(fap_melt,group,by="SampleID")
library(doBy)
fap_mean <- summaryBy(RA~fap+Comparts, fap_merge, FUN = c(mean))
names(fap_mean) <- c('fap', 'Compartments', 'RA')
library(pheatmap)
fap_dcast <- dcast(fap_mean,fap~Compartments,value.var="RA")
rownames(fap_dcast)<-fap_dcast[,1]
fap_dcast<-fap_dcast[,-1]
fap_dcast<-fap_dcast[, c("BS","RS","RE","VE","SE","LE","P")]
scale_test <- apply(fap_dcast, 2, function(x) log2(x+1))
annotation_col = data.frame(Group = factor(rep(c('Belowground','Aboveground'), c(3,4))),row.names = colnames(scale_test)) 
p=pheatmap(scale_test,cluster_row = TRUE, cluster_col = TRUE,cutree_col = 2,filename = paste( "Fap_compartments_heatmap.pdf", sep=""),show_colnames = T,show_rownames = T,annotation_col = annotation_col)

p
```

## b.Boxplot
b, Number of OTUs capable of nitrogen cycling in soil and plant compartments. Boxplot displays the median and interquartile range. Upper and lower whiskers extend to data no more than 1.5× the interquartile range from upper and lower edge of the box, respectively. Letters indicate significant differences among soil and plant compartments (Wilcoxon test, PFDR < 0.05).

This figure is manually modified and stat label added by Adobe Illustrator.
```{r warning=FALSE}
NC = read.table("NC.txt", header=T, row.names=1, sep="\t", comment.char="")
design = read.table("metadata.txt", header=T, row.names=1, sep="\t")
idx = colnames(NC) %in% rownames(design)
NC=NC[,idx]
design=design[colnames(NC),]
NC = as.data.frame(apply(NC, 2, function(x) log2(x+1)))
names(NC)[1] <- c('RA')
library(ggplot2)
library(reshape2)
NC$SampleID=rownames(NC)
NC_merge <-merge(NC,design,by.x = "SampleID",by.y = "row.names", all.x = T)
col=c("#1f78b4","#b2df8a","#a6cee3","#33a02c","#fb9a99","#fdbf6f","#e31a1c")
p = ggplot(NC_merge, aes(x=Comparts, y=RA,color = Comparts)) +
geom_boxplot(position = "dodge", alpha=1, outlier.size=0.3, size=0.5,width = 0.6)+
geom_jitter(size = 1, width = 0.2,alpha=0.5)+
scale_color_manual(values = col)+
  labs(x = 'Compartments', y = 'Number of OTUs with nitrogen cycling (log(x+1))')+scale_x_discrete(limits=c("BS","RS","RE","VE","SE","LE","P"))+
theme(legend.position = "none")+theme(plot.title = element_text(size = 8))+theme_classic() + 
    theme(axis.text.x = element_text(size = 8),axis.text.y = element_text(size = 8))+
    theme(axis.title.y= element_text(size=12))+theme(axis.title.x = element_text(size = 12))+
    theme(legend.title=element_text(size=5),legend.text=element_text(size=5))
p
ggsave("fap_NC_boxplot.pdf", p, width = 150, height = 120,units = "mm")


# A Kruskal-Wallis test is the nonparametric equivalent of the one-way ANOVA and is typically used when the normality assumption is violated. 
model = kruskal.test(RA ~ Comparts, data=NC_merge)
model

NC_merge$Comparts <- factor(NC_merge$Comparts)
group <- levels(NC_merge$Comparts)
group1 <- NULL
group2 <- NULL
p <- NULL

for (i in 1:(length(group) - 1)) {
    for (j in (i + 1):length(group)) {
        group1 <- c(group1, group[i])
        group2 <- c(group2, group[j])
        group_ij <- subset(NC_merge, Comparts %in% c(group[i], group[j]))
        group_ij$Comparts <- factor(group_ij$Comparts, levels = c(group[i], group[j]))
        
        wilcox_test <- wilcox.test(RA ~ Comparts, data = group_ij, alternative = 'two.sided', conf.level = 0.95)
        p <- c(p, wilcox_test$p.value)
    }
}
result <- data.frame(group1, group2, p)
result$padj <- p.adjust(result$p, method = 'BH')	#Benjamini  p 
result
```


## c.Boxplot
 c, Ratio of nifH gene to 16S rRNA in each compartment under fertilisation treatments, reflecting the relative N-fixation potential of bacterial communities (Wilcoxon test, PFDR < 0.05). 

```{r Ratio of nifH tO 16S rRNA, warning=FALSE}
design = read.table("metadata.txt", header=T, row.names=1, sep="\t")
m="CK"
#m="NPK"
#m="NPKM"
sub_design = subset(design, design$Treatment %in% c(m))
nifH = read.table("nifH.txt", header=T,  row.names=1, sep="\t", comment.char="") 
idx = rownames(sub_design) %in% rownames(NC)
sub_design=sub_design[idx,]
nifH=nifH[rownames(sub_design),]
nifH_merge <-merge(nifH,design,by.x = "SampleID",by.y = "row.names", all.x = T)
library(ggplot2)
col=c("#1F78B4","#B2DF8A","#A6CEE3","#33A02C","#FB9A99","#FDBF6F","#E31A1C")
p = ggplot(nifH_merge, aes(x=Comparts, y=LOG,color = Comparts)) +
     geom_boxplot(position = "dodge", alpha=1, outlier.size=0.3, size=0.5,width = 0.6)+
     geom_jitter(size = 1, width = 0.2,alpha=0.5)+
     scale_color_manual(values = col)+
     labs(x = m, y = 'Ratio of nifH tO 16S rRNA (log transformed)')+scale_x_discrete(limits=c("BS","RS","RE","VE","SE","LE","P"))+
  theme_classic()+theme(axis.text.x = element_text(size = 8),axis.text.y = element_text(size = 8))+
    theme(axis.title.y= element_text(size=12))+theme(axis.title.x = element_text(size = 12))+
    theme(legend.title=element_text(size=5),legend.text=element_text(size=5))
p


nifH_merge$Comparts <- factor(nifH_merge$Comparts)
kruskal.test(nifH~Comparts, data = nifH_merge)

group <- levels(nifH_merge$Comparts)
group1 <- NULL
group2 <- NULL
p <- NULL

for (i in 1:(length(group) - 1)) {
    for (j in (i + 1):length(group)) {
        group1 <- c(group1, group[i])
        group2 <- c(group2, group[j])
        group_ij <- subset(nifH_merge, Comparts%in% c(group[i], group[j]))
        group_ij$Comparts <- factor(group_ij$Comparts, levels = c(group[i], group[j]))
        
        wilcox_test <- wilcox.test(nifH ~ Comparts, data = group_ij, alternative = 'two.sided', conf.level = 0.95)
        p <- c(p, wilcox_test$p.value)
    }
}
result <- data.frame(group1, group2, p)
result$padj <- p.adjust(result$p, method = 'BH')
result
```

